---
title: "Homework 2"
author: "Anonymous"
date: 'Due: Thursday May 12th 2022 by 2:00 pm'
output: rmdformats::material
subtitle: 'STAT 363'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggthemes)
library(dplyr)
library(tidyr)
library(tidyverse)
library(maps)
library(mapdata)
```

# General Information

## Logistics

-  Make a copy of this template to write out your solution, DO NOT rename it. Then knit it as `HW02.Rmd`.
- Inside this .Rmd file do not include any personal identifier (such as your name, Odin ID, etc.). 
- Knit your `Rmd` file as `html` and upload both the `Rmd` and `html` files with your solution to D2L in `Assignments > Homework 2` before Thursday May 12th 2022 by 2:00 pm.


## Objectives for this week's homework

1. Practice making figures in both base R and ggplot2
2. Generate hypotheses and explore them graphically



# Your Task

Write down ONLY ONE question or hypothesis you are interested in exploring from the *world powerplant dataset* found in the [information is beautiful](https://informationisbeautiful.net/visualizations/mega-what-the-worlds-biggest-and-most-notable-power-plants/) website, which includes some nice visualizations.  The original data can be found [here](http://datasets.wri.org/dataset/globalpowerplantdatabase) along with a brief description.  You might want to work with just a subset of the data (e.g. one particular country), or you may choose to transform or summarize your data in the way that best allows you to explore your question.

Note that inside of the `globalpowerplant` folder in Canvas the homework folder there are additional documents with information that may prove useful, in particular be sure to look at the `README.txt` file, which includes the metadata (a description of the variables) in the **Fields** section.



**The idea is for you to investigate your question/hypothesis graphically.  To do so, generate the figures you consider necessary -- at least two and at most 4, and note that more figures is not necessarily better, fewer but very well thought-out and informative figures can usually do the trick. <span style="color:orange">At least one of your figures should be built using `base R` and at least one using `ggplot`.</span>**

All titles, axes, and legends should be labeled clearly (no raw variable names). You must have at least one graph with `facet_wrap` or `facet_grid`. You must include at least one manually specified legend. You can use other `geoms` like histograms, bar charts, add vertical or horizontal lines, etc. [You may find this data visualization cheat sheet helpful](https://www.rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf).

 

Your document should be aesthetically pleasing, with some organization. You must write up your observations in words as well as showing the graphs. Use chunk options `echo` and `results` to limit the code/output you show in the `.html`.

Below I load in the data, but be careful, there are many rows that you may need to filter out to remove `NA`s in some of your variables of interest. 

Have fun and Good Luck!!
```{r load_data}
power <- read.csv('globalpowerplant/global_power_plant_database.csv',
                  na.strings = c(NA,"","NA"),
                  header=T)
```

Recall that you can test (and filter out) for `NAs` using the function `is.na`, which is vectorized! If, for example you want to remove rows with NA's in the variable `year_of_capacity_data`, you can use the following line of code  (look at week 2's class notes for more information about this).

```{r examplaNAs}
power <- power %>% filter(!is.na(year_of_capacity_data) )
```

# Your question or hypothesis

Compare three different continents (Asia, Europe, North America), the distribution and comparison of energy in each continent, and use the chart to determine which continent uses more green energy.

# Your base R plot(s)

```{r plot}
#CSV file
df <- read.csv("C:/Users/tulou/Desktop/PSU_Classes/Statistic/STAT_363/Assignment/Homework02/Homework02/globalpowerplant/global_power_plant_database.csv",
               na.strings = c(NA, "", "NA"), 
               header = T)

#map
mapWorld <- map_data("world")
#view(mapWorld)

#Select country to primary_fuel and estimated_generation_gwh
df <- df %>% 
  select(country:primary_fuel, estimated_generation_gwh)
#view(df)

Asia.target <- c("CHN", "TWN", "JPN", "KOR", "IND", "AUS")
test.asian.country <- df %>% 
  filter(country %in% Asia.target)
#test.asian.country

nAmerica.target <- c("MEX", "USA", "CAN")

test.nAmerica.country <- df %>% 
  filter(country %in% nAmerica.target)
#test.nAmerica.country

Europe.target <- c("GBR", "FRA", "DEU", "NOR", "SWE", "FIN", "ESP", "PRT")

test.Europe.country <- df %>% 
  filter(country %in% Europe.target)


continents.summarise <- function(cont){
  cont[is.na(cont)] = 0
  continents <- cont %>% 
    select(primary_fuel, estimated_generation_gwh) %>% 
    group_by(primary_fuel) %>% 
    summarise(total = sum(estimated_generation_gwh))
  return(continents)
}
continents.summarise(test.asian.country)
continents.summarise(test.nAmerica.country)
continents.summarise(test.Europe.country)

Asian.y <- c("Biomass", "Coal", "Gas", "Geothermal", "Hydro ", "Nuclear",
         "Oil", "Solar", "Waste", "Wind")
Asian.X <-c(28811, 5047037, 778603, 2702, 1172734, 333363, 177419, 66058, 7291, 202003)

pie(Asian.X, Asian.y,
    main = "Energies Pie plot in Asia",
    col = rainbow(11))



nAmerica.X <- c(46313, 583576, 392402, 9120, 450532, 117355, 69923,
                5487, 9595, 33282)
nAmerica.y <- c("Biomass", "Coal", "Gas", "Geothermal", "Hydro ", "Nuclear", "Oil", "Other", "Solar", "Wind")

pie(nAmerica.X, nAmerica.y,
    main = "Energies Pie plot in North America",
    col = rainbow(10))

Europe.x <- c(96333, 468998, 238320, 205, 376125, 743113, 23791, 2313, 60315,
              19402, 185302)
Europe.y <- c("Biomass", "Coal", "Gas", "Geothermal", "Hydro ", "Nuclear", "Oil", "Other", "Waste", "Solar", "Wind")

pie(Europe.x, Europe.y,
    main = "Energies Pie plot in Europe",
    col = rainbow(length(Europe.x)))
```

# Your `ggplot` plot(s)

```{r ggplot}
#Asian countries, Japan, Korea, Taiwan, India, Taiwan, & Asutralia
Asia.country.plot <- ggplot() +
  geom_polygon(data = mapWorld, aes(long, lat, group = group)) +
  geom_point(data = test.asian.country,
             aes(longitude, latitude,
                 fill = primary_fuel, group = primary_fuel, color = primary_fuel)) +
  labs(x = "Longitude", 
       y = "Latitude") +
  ggtitle("Asian energy distribution map") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  coord_quickmap()

Asia.country.plot

Asia.ene.plot <- ggplot(test.asian.country, aes(x = primary_fuel, y = estimated_generation_gwh / 10000)) +
  geom_bar(stat = "identity") +
  ylim(0, 550) +
  ggtitle("Histogram of North America Energies") +
  theme(axis.text.x=element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  labs(x = "Primary Fuel",
       y = "Estimated Generation gwh(* 10000)") +
  theme(axis.text.x=element_text(angle = 45, hjust = 1))
Asia.ene.plot

#North America courties, America, Canada, Mexico
nAmerica.country.plot <- ggplot() +
  geom_polygon(data = mapWorld, aes(long, lat, group = group)) +
  geom_point(data = test.nAmerica.country,
             aes(longitude, latitude,
                 fill = primary_fuel, group = primary_fuel, color = primary_fuel)) +
  labs(x = "Longitude", 
       y = "Latitude") +
  ggtitle("North America energy distribution map") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  coord_quickmap()

nAmerica.country.plot

nAmerica.ene.plot <- ggplot(test.nAmerica.country, aes(x = primary_fuel, y = estimated_generation_gwh / 10000)) +
  geom_bar(stat = "identity") + 
  ylim(0, 60) +
  ggtitle("Histogram of North America Energies") +
  theme(axis.text.x=element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  labs(x = "Primary Fuel",
       y = "Estimated Generation gwh(* 10000)")
nAmerica.ene.plot

#Europe countries, United Kingdom, Franch, Germany, Norway, Sweden, Finland, Spain, Portugal
Europe.country.plot <- ggplot() +
  geom_polygon(data = mapWorld, aes(long, lat, group = group)) +
  geom_point(data = test.Europe.country,
             aes(longitude, latitude, 
                 fill = primary_fuel, group = primary_fuel, color = primary_fuel)) +
  labs(x = "Longitude", 
       y = "Latitude") +
  ggtitle("European energy distribution map") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  coord_quickmap()

Europe.country.plot

Europe.ene.plot <- ggplot(test.Europe.country, aes(x = primary_fuel, y = estimated_generation_gwh / 10000)) +
  geom_bar(stat = "identity") +
  ylim(0, 80) +
  ggtitle("Histogram of Europe Energies") +
  theme(axis.text.x=element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  labs(x = "Primary Fuel",
       y = "Estimated Generation gwh(* 10000)") 
Europe.ene.plot
```



